### 2016-09-13 paper/old_sections.tex
stashing some paper changes

\if{false} %--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--
%------------------------------------------------------------------------------------------------------------------------%
\subsection{Constant vs Field Parameters} \label{sec:constvfield}
%------------------------------------------------------------------------------------------------------------------------%
It is not necessary for the low- and high-fidelity models to differ in the physics included. A low-fidelity model may be more computationally tractable due to a reduced number of degrees of freedom rather than reduced nonlinearity. In this section, we consider two models which differ in the space to which the parameter belongs, with the low-fidelity model having fewer degrees of freedom.
%------------------------------------------------------------%
\subsubsection{Problem Setup}
%------------------------------------------------------------%
We consider the same high-fidelity model as in \cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U,
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in\R
\end{equation}
differs from the high-fidelity model only in that the parameter $q$ is a constant instead of a field. Then the intermediate mixed-fidelity models have parameter fields which are non-constant in only portions of the domain. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. The domain, mesh, boundary conditions, and velocity field, as well as the observations, unknown parameters to be inferred, and QoI, remain the same as described in \cref{sec:cdvcdr}. As the inverse problem is ill-posed, except for perhaps in the case where the low-fidelity model is used throughout the domain, regularization is added; the Tikhonov regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient.

@@

In other words, on using \cref{alg:refSeries} for a highly nonlinear problem, even with the worst case linear solver, one can get to $<1\%$ error in the QoI, with a $20\%$ reduction in computational cost, while avoiding any need for continuation to handle nonlinearities. }
\fi{false} %--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--*--

>>>>

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Complexity Analysis}\label{sect:alg_complexity}
%------------------------------------------------------------------------------------------------------------------------%
We now derive the computational complexity of using Algorithm~\cref{alg:refSeries} to solve an inference problem, and compare it to extant strategies. Let the state, adjoint, and parameter variables each have $N$ degrees of freedom, and assume that an `all at once' strategy is used to solve the KKT system. The cost of single linear solve with such a system is $\orderof{3N}^{\gamma}$, where $\gamma$ depends on the linear solver used~\cite{}. The cost of solving for the auxiliary variables, a linear system, is $\orderof{3N}^{\gamma}$. \red{The parts of the supplementary adjoint corresponding to the primary and auxiliary variables can be solved separately, and thus the cost of computing the supplementary adjoint is also $\orderof{3N}^{\gamma}$.}

If one utilizes Algorithm~\cref{alg:refSeries} to solve the inference problem, with $T$ adaptive iterations the total computational cost is,
%
\begin{align}
\label{eq:cost_adapt}
C_{MF} &= \sum_{i=1}^{T}L_{i}(3N)^{\gamma} + 1(3N)^{\gamma} + 2(3N)^{\gamma} \nonumber \\
&= \sum_{i=1}^{T} (L_i + 3) (3N)^{\gamma}
\end{align}
% 
where $L_i$ is the number of nonlinear solver steps needed to solve the KKT system for the low or mixed fidelity model at each adaptive step. The corresponding cost for solving the high fidelity inference problem, with a nonadaptive algorithm, based on a continuation scheme is,
%
\begin{equation}
\label{eq:cost_nadapt}
C_{HF} = \sum_{i=1}^{C}K_{i}(3N)^{\gamma}
\end{equation}
% 
where $K_i$ is the number of nonlinear solver steps needed to solve the KKT system for the high fidelity model at each continuation step, and $C$ is the number of continuation steps.

Comparing Eq.~\eqref{eq:cost_adapt} and~\eqref{eq:cost_nadapt}, we have,
%
\begin{equation}
\label{eq:cost_compare_0}
C_{MF} < C_{HF} \implies \sum_{i=1}^{T} (L_i + 3) < \sum\limits_{i=1}^{C}K_{i}
\end{equation}
%
If the low and multi-fidelity models are linear, or nonlinear in a very small region, then $L_i \approx 1$, and we have the following condition,
%
\begin{equation}
\label{eq:cost_compare_1}
T < \frac{\sum\limits_{i=1}^{C}K_{i}}{4},
\end{equation}
% 
for the adaptive algorithm to be less costly than solving the high-fidelity inverse problem. \red{I'm not sure this section makes sense, since this doesn't seem to hold for any of the actual runs we've done...our mixed models take fewer nonlinear steps than the HF model, but not close enough to one...}

%
%
%
%
%
%
%
%

\red{\Cref{alg:refSeries} can be naturally extended to an online-offline setting, where a multi-fidelity model adaptive constraint can be generated for a given data set in the offline phase. In the online phase, additional data points are added to the data set, and instead of solving the complete inverse problem with all the data points, and the high-fidelity constraints, an iterative algorithm can be initiated using the multi-fidelity representation developed in the offline phase. The adaptive procedure, with the supplementary adjoint can be used to adapt the preexisting multi-fidelity representation to the new data points. } 
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adapt a pre-computed mixed-fidelity model to new data.}
\label{alg:refonoff}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Given a data set $d_1 \in \R^{n_{d_1}}$, and an adaptively build a mixed-fidelity model for low error in the QoI, use \cref{alg:refSeries} to build a MF model MF$_1$} 
\Procedure{$\texttt{Online}$}{HF model, MF model MF$_1$, \texttt{errTol}, $d_2 \in \R^{n_{d_2}}$}
	\State{$i\gets1$}
	\State{Solve for stationary point $\Psi_{MF_1}$ of augmented Lagrangian $\mathcal{M}_{MF_1}$ formed using new data set $d_2$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_1}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_1$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

	$\epsilon_0=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF_1})(\Lambda_1)+\mathcal M_{HF}(\Psi_{MF_1})-\mathcal M_{MF_1}(\Psi_{MF_1})$}
	\State{Calculate QoI $I(q_{MF_1},u_{MF_1})$}
	\While{$i<$ \texttt{maxIter} and $|\epsilon_i/I(q_{MF_i},u_{MF_i})|>$ \texttt{errTol}}
		\State{\begin{varwidth}[t]{\linewidth}Localize $\epsilon_i$ (see \cref{sec:errLocal}) and use this decomposition to guide \par\hskip\algorithmicindent formation of new mixed-fidelity model MF$_{i+1}$\end{varwidth}}
		\State{$i\gets i+1$}
		\State{Solve for stationary point $\Psi_{MF_i}$ of augmented Lagrangian $\mathcal{M}_{MF_i}$}
		\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_i}$, for

		$\quad\quad$supplementary adjoint $\Lambda_i$ (see \cref{eq:superAdjEq})}
		\State{Compute QoI error estimate

		$\quad\quad \epsilon_i=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF_i})(\Lambda_i)+\mathcal M_{HF}(\Psi_{MF_i})-\mathcal M_{MF_i}(\Psi_{MF_i})$}
		\State{Calculate QoI $I(q_{MF_i},u_{MF_i})$}
	\EndWhile \\
\Return{QoI estimate $I(q_{MF_i},u_{MF_i})$ for the new data set $d_2$}
\EndProcedure
\Statex
\end{algorithmic}
\end{algorithm}
%
\Cref{alg:refonoff} lends itself to a quasi-incremental data assimilation approach, where the data set $d_2$ is an augmentation of the original set $d_1$. In such a situation, we expect that the model refinement will be concentrated around the new data points, that also inform the QoI. Thus, in the online phase, even though the inverse problem will be solved with all data points, the use of high-fidelity models will be limited, and the bulk of the computation (computing the auxilary variables, and supplementary adjoint) will be focused on identifying those new data points, which have the maximum impact on the QoI.

%
%
%
%
%
%
%
%
%
%

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Constant vs Field Parameters} \label{sec:constvfield}
%------------------------------------------------------------------------------------------------------------------------%
It is not necessary for the low- and high-fidelity models to differ in the physics included. A low-fidelity model may be more computationally tractable due to a reduced number of degrees of freedom rather than reduced nonlinearity. In this section, we consider two models which differ in the space to which the parameter belongs, with the low-fidelity model having fewer degrees of freedom.
%------------------------------------------------------------%
\subsubsection{Problem Setup}
%------------------------------------------------------------%
We consider the same high-fidelity model as in \cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U,
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in\R
\end{equation}
differs from the high-fidelity model only in that the parameter $q$ is a constant instead of a field. Then the intermediate mixed-fidelity models have parameter fields which are non-constant in only portions of the domain. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. The domain, mesh, boundary conditions, and velocity field, as well as the observations, unknown parameters to be inferred, and QoI, remain the same as described in \cref{sec:cdvcdr}. As the inverse problem is ill-posed, except for perhaps in the case where the low-fidelity model is used throughout the domain, regularization is added; the Tikhonov regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient.

@@

In other words, on using \cref{alg:refSeries} for a highly nonlinear problem, even with the worst case linear solver, one can get to $<1\%$ error in the QoI, with a $20\%$ reduction in computational cost, while avoiding any need for continuation to handle nonlinearities. }

%
%
%
%
%
%
%
%
